Efficient spherical Convolutional Neural Networks with Healpix sampling for cosmological applications

### 2018-05-14 scnn.tex
small changes

In an attempt to get the best of the two approaches, we follow a third
direction. We use a graph to adapt to the particular structure of the sphere.
While the convolution remains efficient, it still captures well the spherical
structure and is particularly adapted to the HealPix sampling.

@@

As presented in~\cite{cohen2018spherical}, the most mathematical approach to
extend the convolution on a sphere is to use a spherical Fourier transform. The
convolution is then simply defined as the product in the spectral domain. This
approach requires one Fourier and one inverse Fourier transform per convolution
which remain, even with accelerated algorithms, expensive. For 2-dimensional
images, an efficient convolution can be achieved when the convolution kernel is
localized (for example a 5x5 pixel patch) by doing the computation directly in
the signal domain. Unfortunately, this approach cannot be directly extended to
the spherical case. Hence, the main idea of this contribution is to leverage
graph signal processing~\cite{shuman2013emerging} to define a spherical
convolution that can be computed directly in the signal domain.

@@

Because of the irregular sampling, there is no obvious way to define the convolution
directly from the pixel values. Hence we first define a Fourier transform on the
graph.

@@

\paragraph{Example: heat diffusion}
Let us consider the heat diffusion problem 
\begin{equation} \label{eq:heat_equation}
\L \b{f}(t) = \tau \partial_t \b{f}(t),
\end{equation}
where $\b{f}(t): \mathbb{R}_+ \rightarrow \mathbb{R}^N$. Given the initial condition 
$\b{f}(0)$, the solution of~\ref{eq:heat_equation} can be expressed as
\begin{equation}
\b{f}(t) = e^{-\L \tau t} \b{f}(0) = \U e^{-\bLambda t \tau} \U^* \g{f}(0) = K_t(\L) \b{f}(0), 
\end{equation}
which is a convolution of the signal $\b{f}(0)$ with the kernel $K_t(x)=e^{-\tau
t x}$. Here the kernel $K_t$ is applied to the graph eigenvalues $\bLambda$ that
correspond to the squared frequencies and hence $K_t$ can be considered a
Gaussian kernel. In Figure~\ref{fig:gaussian_filters_comparizon} top, we show
the effect of the convolution by diffusing a unit of heat for $\tau=1$ and various 
time $t$. In  Figure~\ref{fig:gaussian_filters_visualization} of
Appendix~\ref{app:filter_visualization}, we show different visualizations of the
convolution kernel $K_t$. We also compare the graph convolution with the
spherical symmetric Gaussian smoothing
(Figure~\ref{fig:gaussian_filters_comparizon} bottom) and observe that both
techniques lead to similar results. While the graph convolution is different
from a true spherical convolution, it can, providing the correct
parameterization, well approximate it.

@@

It is another sign that the constructed graph is able to capture the spherical
structure of the HealPix sampling. Furthermore, in some application, where the
exactitude of the convolution is not a requirement, such as de-noising, it means
that we could use the graph convolution instead. Depending on the convolution kernel
this may reveal much more efficient. In the case of a neural network, the fact
that both convolution are not identical is not very relevant since the network
will adapt the convolution at use.

@@

We would like to build a robust classifier, i.e. an automatic way to
discriminate between the two classes with the presence of noise. For all the
experiments, we work with zero mean Gaussian noise. \todo{Tomek: could you put a
justification.} All classifiers are trained with noisy samples. To improve the
stability of the spherical CNN, we slowly increase the amount of noise added to
the training data over the iterations. Furthermore in order to avoid
overfitting, the noise is randomly produced at each iteration. We trained our
model between $5$ and $40$ epochs depending on the complexity and the size of
the problem. The testing set was simply used to assess the global performance of
the network and to assess if were overfitting. 
Training took between $1$ and $10$ hours using a Nvidia 1080 TI.

We compare the spherical CNN with a handcrafted feature classifier. \cite{...}
\todo{Tomek: we probably need 1 or two other classifiers.} The features are
constructed using a normalized histogram of the values. We then train a linear
SVN classifier. We tried different kernels, but they were not leading to better
results. For a fair comparison, we increase the amount of training data of the
SVN classifier until the testing error plateau to fix value.

>>>>

In an attempt to get the best of the two first approaches, we follow a third
direction. We use a graph to adapt to the particular structure of the sphere.
While the convolution remains efficient, it still captures well the spherical
structure and is particularly adapted to the HealPix sampling.

@@

As presented by~\cite{cohen2018spherical}, the most mathematical approach to
extend the convolution on a sphere is to use a spherical Fourier transform. The
convolution is then simply defined as the product in the spectral domain. This
approach requires one Fourier and one inverse Fourier transform per convolution
which remain, even with accelerated algorithms, expensive. For 2-dimensional
images, an efficient convolution can be achieved when the convolution kernel is
localized (for example a 5x5 pixel patch) by doing the computation directly in
the signal domain. Unfortunately, this approach cannot be directly extended to
the spherical case. Hence, the main idea of this contribution is to leverage
graph signal processing~\cite{shuman2013emerging} to define a spherical
convolution that can be computed (and back-propagated) directly in the signal domain.

@@

Because of the irregular sampling, there is no obvious way to define the convolution
directly from the pixel values. Hence we will first make a detour and define a Fourier transform on the
graph.

@@

\paragraph{Example: heat diffusion}
Let us consider the heat diffusion problem 
\begin{equation} \label{eq:heat_equation}
\L \b{f}(t) = \tau \partial_t \b{f}(t),
\end{equation}
where $\b{f}(t): \mathbb{R}_+ \rightarrow \mathbb{R}^N$. Given the initial condition 
$\b{f}(0)$, the solution of~\ref{eq:heat_equation} can be expressed as
\begin{equation}
\b{f}(t) = e^{-\L \tau t} \b{f}(0) = \U e^{-\bLambda t \tau} \U^* \g{f}(0) = K_t(\L) \b{f}(0), 
\end{equation}
which is a convolution of the signal $\b{f}(0)$ with the kernel $K_t(x)=e^{-\tau
t x}$. Here the kernel $K_t$ is applied to the graph eigenvalues $\bLambda$ that
are one generalization of the squared frequencies on the graph. Hence the kernel $K_t$ is also a generalization of the
Gaussian kernel on the sphere. In Figure~\ref{fig:gaussian_filters_comparizon} top, we show
the effect of the convolution by diffusing a unit of heat for $\tau=1$ and various 
time $t$. In  Figure~\ref{fig:gaussian_filters_visualization} of
Appendix~\ref{app:filter_visualization}, we show different visualizations of the
convolution kernel $K_t$. We also compare the graph convolution with the
spherical symmetric Gaussian smoothing
(Figure~\ref{fig:gaussian_filters_comparizon} bottom) and observe that both
techniques lead to similar results. While the graph convolution is different
from a true spherical convolution, it can, providing the correct
parameterization, well approximate it.

@@

This result is another sign that the constructed graph is able to capture the spherical
structure of the HealPix sampling. Furthermore, in some application, where the
exactitude of the convolution is not a requirement, such as de-noising, it means
that we could use the graph convolution instead. Depending on the convolution kernel
this may reveal much more efficient. In the case of a neural network, the fact
that both convolution are not identical is not very relevant since the network
will adapt the convolution at use.

@@

We would like to build a robust classifier, i.e. an automatic way to
discriminate between the two classes with the presence of noise. For all the
experiments, we work with centered Gaussian noise. \todo{Tomek: could you put a
justification.} All classifiers are trained with noisy samples. To improve the
stability of the spherical CNN, we slowly increase the amount of noise added to
the training data over the iterations. Furthermore in order to avoid
overfitting, the noise is randomly produced during the training process. We trained our
model between $5$ and $40$ epochs depending on the complexity and the size of
the problem. The testing set was simply used to assess the global performance of
the network and to assess if it was overfitting. 
Training took between $1$ and $10$ hours using a Nvidia 1080 TI.

We compare the spherical CNN with a handcrafted feature classifier. \cite{...}
\todo{Tomek: we probably need 1 or two other classifiers.} The features are
constructed using a normalized histogram of the values. We then train a linear
SVN classifier. We tried different kernels, but they were not leading to better
results. For a fair comparison, we increase the amount of training data of the
SVN classifier until the testing error plateau to a fix value.
